Informative Autonomous Sampling of Oceanographic Variables Using Joint Excursion Sets

### 2019-10-09 main_new.tex
Minor changes & rephrased parts around sensitivity on simulation results.

\input{commands.tex}
\usepackage{amsmath}
\usepackage{pstricks,pst-grad}
\usepackage{graphicx}
\usepackage{floatrow}
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\floatsetup[table]{capposition=top}
\usepackage{subfigure}
\usepackage[utf8]{inputenc}
\usepackage{booktabs}                     % horizontal lines in tables

@@

\begin{abstract}
  Improving and optimizing oceanographic sampling is a crucial task
  for marine science and maritime management. Faced with limited
  resources to understand dynamic processes in the water-column, the
  combination of statistics and autonomous robotics provide new
  opportunities for studying experimental design. In this work we develop
  methods for efficient spatial sampling applied to the mapping of
  coastal processes by providing informative descriptions of spatial
  characteristics of ocean phenomena. Specifically, we define a design
  criterion based on reducing uncertainty in the joint excursions of
  random fields, and derive analytical expressions for the expected
  Bernoulli variance reduction in the excursion set of spatial
  temperature and salinity variables, providing a description of
the boundary uncertainty that can be used prioritize sampling efforts at locations that are
ambiguous, making exploration more effective. We show how these can be
  applicable in static and sequential designs, building on myopic and
  non-myopic sampling strategies. We use simulations to study the
  properties and differences between state-of-the-art approaches,
  followed by results from field deployments with an autonomous
  underwater vehicle as part of a case study mapping the boundary of a
  river plume. The results demonstrate the potential of combining statistical methods and robotic platforms, and the abilities of the derived methods to effectively inform and execute data-driven environmental sampling.

@@

Our objective is to leverage statistical tools to characterize a river
plume, focusing on spatial separation of cold freshwater from the
river and warmer saline waters of the fjord. ES and EPs are useful
starting points to measure the ability to classify water masses. The
goal of experimental sampling is to improve this
characterization. %, as quantified via changes in the EPs.

@@

The derived closed-form solutions for the expected IBV are still
important building blocks when constructing the sampling designs, as
they will be used to score the different adaptive designs. % Efficient
% calculation is important for using adaptive survey designs with
% robotic platforms.

@@

We use a bivariate GP model for temperature and salinity, where we
specify the prior mean:
\begin{equation}\label{m}
    \bmu(\bx)=E 
    \begin{bmatrix}
    \xi_T(\bx) \\
    \xi_S(\bx) 
    \end{bmatrix}=\begin{bmatrix} \mu_{T}(\bx)\\
\mu_{S}(\bx)
\end{bmatrix} 
= \begin{bmatrix} \beta_{T,0} + \beta_{T,1} x_{\mbox{West}} \\
\beta_{T,1} + \beta_{S,1} x_{\mbox{West}}
\end{bmatrix}.
\end{equation}
This situation
mimics that of a river mouth opening from the south, with the water
masses pulled to the east. The regression parameters are  $\beta_{T,1}=0.065$ and
$\beta_{S,1}=0.1$ for the slope, with west coordinate and $\beta_{T,0}=5.8$
and $\beta_{S,1}=29.0$ for the intercept at a reference map coordinate.

@@

For a fixed survey length, the expected IBV is analytically available
\kc{I don't understand what ``analytically available''
  means}. However, for the sequential approaches this is not the
case. For comparison the properties are therefore evaluated using
Monte Carlo sampling over several replicates of realizations from the
prior while conducting simulated sequential surveys for each
one. Fig. \ref{fig:Eprob} shows the conditional EP, given data
gathered along the north-south survey lines in these figures. In the
Monte Carlo replicates, such results are repeatedly computed to
approximate the expected IBV. We also compare predictive performance
measured by root mean square error (RMSE) for temperature and salinity
estimates as also the variance reduction in these two variables. It is
important to note that the objective function used by the agent is
focused on reducing the expected IBV, but we nevertheless expect that
we will achieve good predictive performance for criteria such as RMSE
as well. Another non-statistical criteria that is important for
practical purposes is the computational time needed for the strategy,
as this will impact the performance for an embedded system.

@@

Each strategy is conducted on an equilateral grid as shown in
Fig. \ref{fig:wp_graph}. The sequential sampling agent starts at the
center East-West coordinate at the southern end of the domain (node
53). The AUV moves along edges in the waypoint graph while the vehicle
gathers measurements. The data is assimilated into the GP model before
an evaluation of the next node to sample is conducted at the end of
the edge. \kc{Note there are plenty of language errors related to
  tense. Pl. do not mess with any more changes from here onwards.}

@@

In Fig. \ref{fig:route_choices}, the realized sampling paths for each
of the sequential schemes and static designs are shown. The
\textit{naive} strategy often gets stuck in the southern part of the
domain because it is too focused on the probabilities near $0.5$. The
\textit{myopic} strategy covers a wider domain than the naive or
look-ahead. There are several reasons for this. First, a greedy
approach will tend to put more emphasis on locations close to the
agent, which may lead away from the centre. Second, as the agent
evaluates the impact of locations further away (look-ahead) where
assimilated data has less predictive power, the GP prior model (which
is centered here) will act to restrict paths deviating from the
central zone.

%We studied the sensitivity of the results by modifying the input parameters to different prior correlations between temperature and salinity, standard deviations, and spatial correlation range.  In all runs,  the \textit{myopic} and \textit{look-ahead} strategies perform the best in terms of IBV, and much better than \textit{naive}. The \textit{look-ahead} strategy seems to be substantially better than the \textit{myopic} only for very small prior standard deviation or very large spatial correlation range. The \textit{static\_north} design continues to be the best static design for IBV, while \textit{static\_zigzag} is the best design for the other predictive performance measures, especially so with large spatial correlation range. We also ran simulation studies with only temperature data, and for realistic correlation levels between temperature and salinity the effects are not much worse in terms of IBV when only temperature data are available. In addition to the comparison made in Table \ref{tab:sim_rhoab}, the current setting includes spatial correlation and this likely reduces the additional influence of having bivariate data. However, it seems that having temperature data alone does a substantially worse job in terms of explained variance

\subsection{Performance Metrics}

@@

\newpage

>>>>

\input{commands.tex}
\usepackage{amsmath}
\usepackage{pstricks,pst-grad}
\usepackage{graphicx}
\usepackage{floatrow}
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\floatsetup[table]{capposition=top}
\usepackage{subfigure}
\usepackage[utf8]{inputenc}
\usepackage{booktabs}                     % horizontal lines in tables
\usepackage{comment}

@@

\begin{abstract}
  Improving and optimizing oceanographic sampling is a crucial task
  for marine science and maritime management. Faced with limited
  resources to understand dynamic processes in the water-column, the
  combination of statistics and autonomous robotics provide new
  opportunities for studying experimental design. In this work we develop
  methods for efficient spatial sampling applied to the mapping of
  coastal processes by providing informative descriptions of spatial
  characteristics of ocean phenomena. Specifically, we define a design
  criterion based on reducing uncertainty in the joint excursions of
  random fields, and derive analytical expressions for the expected
  Bernoulli variance reduction in the excursion set of spatial
  temperature and salinity variables, providing a description of
the boundary uncertainty that can be used to prioritize sampling efforts at locations that are
ambiguous, making exploration more effective. We show how these can be
  applicable in static and sequential designs, building on myopic and
  non-myopic sampling strategies. We use simulations to study the
  properties and differences between state-of-the-art approaches,
  followed by results from field deployments with an autonomous
  underwater vehicle as part of a case study mapping the boundary of a
  river plume. The results demonstrate the potential of combining statistical methods and robotic platforms, and the abilities of the derived methods to effectively inform and execute data-driven environmental sampling.

@@

Our objective is to leverage statistical tools to characterize a river
plume, focusing on spatial separation of cold freshwater from the
river and warmer saline waters of the fjord. ES and EPs are useful
starting points to measure the ability to classify water masses. The
goal of experimental sampling is to improve this
characterization, as quantified via changes in the EPs.

@@

%The derived closed-form solutions for the expected IBV are still important building blocks when constructing the sampling designs, as they will be used to score the different adaptive designs. % Efficient
% calculation is important for using adaptive survey designs with
% robotic platforms.

@@

We use a bivariate GP model for temperature and salinity, where we
specify the prior mean:
\begin{equation}\label{m}
    \bmu(\bx)=E 
    \begin{bmatrix}
    \xi_T(\bx) \\
    \xi_S(\bx) 
    \end{bmatrix}=\begin{bmatrix} \mu_{T}(\bx)\\
\mu_{S}(\bx)
\end{bmatrix} 
= \begin{bmatrix} \beta_{T,0} + \beta_{T,1} x_{\mbox{West}} \\
\beta_{S,0} + \beta_{S,1} x_{\mbox{West}}
\end{bmatrix}.
\end{equation}
This situation
mimics that of a river mouth opening from the south, with the water
masses pulled to the east. The regression parameters are  $\beta_{T,1}=0.065$ and
$\beta_{S,1}=0.1$ for the slopes with west coordinate, and $\beta_{T,0}=5.8$
and $\beta_{S,0}=29.0$ for the intercept at a chosen reference map coordinate.

@@

For a fixed survey length, a closed form expression for the expected IBV is available as in Eq. \eqref{sur}. However, for the sequential approaches this is not the
case. For comparison the properties are therefore evaluated using
Monte Carlo sampling over several replicates of realizations from the
prior while conducting simulated sequential surveys for each
one. Fig. \ref{fig:Eprob} shows the conditional EP, given data
gathered along the north-south survey lines in these figures. In the
Monte Carlo replicates, such results are repeatedly computed to
approximate the expected IBV. We also compare predictive performance
measured by root mean square error (RMSE) for temperature and salinity
estimates as also the variance reduction in these two variables. It is
important to note that the objective function used by the agent is
focused on reducing the expected IBV, but we nevertheless expect that
we will achieve good predictive performance for criteria such as RMSE
as well. Another non-statistical criteria that is important for
practical purposes is the computational time needed for the strategy,
as this will impact the performance for an embedded system.

@@

Each strategy is conducted on an equilateral grid as shown in
Fig. \ref{fig:wp_graph}. The sequential sampling agent starts at the
center East-West coordinate at the southern end of the domain (node
53). The AUV moves along edges in the waypoint graph while the vehicle
gathers measurements. The data is assimilated into the GP model before
an evaluation of the next node to sample is conducted at the end of
the edge.

@@

In Fig. \ref{fig:route_choices}, the realized sampling paths for each
of the sequential schemes and static designs are shown. The
\textit{naive} strategy often gets stuck in the southern part of the
domain because it is too focused on the probabilities near $0.5$. The
\textit{myopic} strategy covers a wider domain than the naive or
look-ahead. There are several reasons for this. First, a greedy
approach will tend to put more emphasis on promising locations close to the
agent, which may lead away from the centre. Second, as the agent
evaluates the impact of locations further away (look-ahead) where
assimilated data has less predictive power, the GP prior model (which
is centered here) will act to restrict paths deviating from the
central zone.

We studied the sensitivity of the results by modifying the input parameters to different prior correlations between temperature and salinity, standard deviations, and spatial correlation range.  In all runs,  the \textit{myopic} and \textit{look-ahead} strategies perform the best in terms of IBV, and much better than \textit{naive}. The \textit{look-ahead} strategy seems to be substantially better than the \textit{myopic} only for very small prior standard deviation or very large spatial correlation range. The \textit{static\_north} design continues to be the best static design for IBV, while \textit{static\_zigzag} is the best design for the other predictive performance measures, especially so with large spatial correlation range. We also ran simulation studies with only temperature data, and for realistic correlation levels between temperature and salinity the effects are not much worse in terms of IBV when only temperature data are available. In addition to the comparison made in Table \ref{tab:sim_rhoab}, the current setting includes spatial correlation and this likely reduces the additional influence of having bivariate data. However, it seems that having temperature data alone does a substantially worse job in terms of explained variance

\begin{comment}
\subsection{Performance Metrics}

@@

\end{comment}
